???A quantitative framework for investigating the reliability of network construction

### 2017-10-27 manuscript/CI_interactions.tex
spellcheck, all sections now included?

    The main objective of this study is to develop analytical tools that could be used to better represent the uncertainty in the estimation of pairwise interaction probability. We start with the description of a new Bayesian approach to estimate ecological interactions and of a related R package. [[We first illustrate the framework with a simple quantitative example. Next, we use the framework to provide analytical criteria for improving the sampling of rare interations. Finally, we discuss the implications of uncertainty in the estimation of pairwise interactions on the properties of ecological networks.]] Through these efforts, we demonstrate both the utility of our approach and the importance of acknowledging the uncertainty inherent in ecological networks.
    % List of sections may not be accurate any more. Update when paper reaches its final form.

@@

THIS SECTIONS BUILDS ON THE SALIX DATA TO ILLUSTRATE THE GAPS THERE ARE IN THE METHOD. DESPITE THE INTENSITY OF THE SAMPLING, THERE ARE SEVERAL HOLES.

\section*{A naive qunatification of interaction probability and uncertainty}

@@

  As a start, we will consider how we could naively quantify intetraction probability and its associated uncertainty for an interaction that has not yet been observed. We consider that a pair of species have been observed $n$ times co-occurring at one location, and the number of times they have been observed interacting is $k = 0$. We aim to evaluate the uncertainty of this interaction. The occurrence of an interaction is a Bernoulli trial. Consequently, the number of success $k$ over $n$ trials will follow a binomial distribution:

@@

  The variance of a Bernoulli experiment is simply $n\lambda$(1-$\lambda$). It is important to remember this variance reports the variability of the number of success $k$ for $n$ trials and is not the variance associated with the estimation of $\lambda$.
  Given this variance, it is possible to compute the confidence interval for the MLE of $\lambda$ using any of several methods, including the \emph{Wilson score interval}, the \emph{Clopper-Pearson interval}, and the \emph{Agresti-Coull Interval}. Finding this estimate is therefore quite straightforward, but it nonetheless has two drawbacks. First, $\lambda$ is a random variable with a given uncertainty and whose distribution is unknown, it is not a single point estimate. This means that even if $k = 0$ in a given sample, this does not necessarily imply that the two species will never interact. It means that no interaction is the most likely situation but there is nonetheless a significant chance that the two species could interaction. In contrast with the situation where $k>0$, when we are sure that an interaction is feasible, 
  % (in which case, the ratio $k/n$ is smaller than $n$ because of detection and process uncertainty),  % This parenthetical makes the sentence very long. Is it necessary?
  we are not sure that absence of interactions is caused by $L=0$. Second, and most importantly, the number of trials $n$ might be very low in many instances (and some pairs of species might even not be documented once). Consequently, the uncertainty in the estimation of $\lambda$ might be considerable. We illustrated in Fig. X how the estimate of $\lambda$ could vary with sampling size using the Clopper-Pearson interval. It shows that the 95\% confidence interval requires a signifcant amount of sampling in order to be sure that species are not interacting. 
  % Is the Clopper-Pearson interval the one in Dom's code snippets? If so, this is Fig.~\ref{Salix_cdfs}

@@

  What does this question mean exactly? We consider that there are three nested levels of uncertainty making the occurance of any interaction a stochastic process:

@@

    It could be that the interaction is feasible, i.e. $L=1$, but that it does not occur at a given location or moment in time because there are local constraints preventing it from occurring. We define the realization of the interaction process $X$ given that the interaction is feasible as a stochastic process with associated probability $P(X|L)=\chi$. This phenomenon of interaction contingencies is usually not considered in network studies, but there is a rich literature in community ecology about the contingencies of interactions~\citep{}. We imagine for instance that an interaction between a gall and a parasitoid might not be recorded at a location because it was the wrong time of the year and the gall had not yet formed (phenological constraints), a late frost killed the larvae (an abiotic environmental constraint), another parasitoid species competitively excluded the parasitoid species of interst (a biotic environmental constraint) or simply the parasitoid is too rare to interact with all of its viable hosts (an abundance constraint).

@@

    Lastly, measurement errors are always source of uncertainty in the observation of ecological processes. We define the detection of an interaction $D$, given an interaction is feasible and occurs in the local conditions, as a stochastic process with associated probability $P(D|X,L)=\delta$. Detection failure could happen for several reasons, for instance if the rearing of a parasitoid fails with inappropriate lab conditions or because of species mis-identification. Some sources of detection error could be minimized with appropriate sampling effort ($\delta$ will converge to one with increasing number of collected galls), but other sources are often difficult to reduce (e.g. the occurrence of cryptic species might require molecular analysis for appropriate taxonomic identification).

@@

  For the ecologist working on collecting network data, and on later interpreting these data, the considerations above raise a major challenge: how may we infer whether unobserved interactions went undetected due to sampling, or whether they truly do not occur. How then may we refine our sampling approaches to reduce uncertainties, and to gain insights into the impact of multiple processes on field observations? Importantly, some sources can be minimized with appropriate sampling design and efforts, while other sources are difficult or impossible to reduce – since they are generated by chance variation created by the very process in which we are interested.

  The basic information that we are collecting consists of matrices of interaction partners. Our count of interactions between species A and species B reflects uncertainty with respect to whether two species interact when co-occurring at a given site, chance variation in whether the interaction is realized, and uncertainty with respect to whether it is detected when realized. In this case, what can we do to pin down any of these different types of variation?

@@

  \begin{figure}
    \caption{\textbf{A-B)} Most pairs of \emph{Salix} and galler or galler and parasitoid species were never observed co-occurring despite the high levels of replication in our example dataset. For those that were observed together at least once, the number of observed co-occurances was generally small (\textless10). Here we show a histogram of the number of pairs of species observed co-occurring at least once. 3986 pairs of species were never observed co-occurring: these pairs are omitted from the histogram. \textbf{C-D)} Most pairs of species that were observed at the same site were never observed interacting. Here we show a histogram of the number of observed interactions within pairs of co-occurring species. \textbf{E-F)} Here we show, for each species pair, the number of observed interactions plotted against the number of observed co-occurances. Most pairs either are never observed interacting or interact almost every time they co-occur. The red, dashed line indicates a 1:1 relationship between interactions and co-occurances.}
    \label{histograms}
    \includegraphics*[width=.8\textwidth]{figures/Salix_Galler_histogram.eps}
    \end{figure}

@@

  As a solution towards deciding whether unobserved interactions were undetected due to sampling, or whether they truly do not occur,~\citet{Weinstein2017} used daily repeated sampling rounds to estimate the daily probability of detecting a hummingbird interaction, and to thereby model the observation and process mechanisms~\citep{Weinstein2017}. While conceptually attractive, this approach is unsuitable for interactions occurring over longer time scales (e.g., associations between hosts and parasitoids with a single generation per year), or very rare interactions (since no matter how frequently and extensively we sample, there will still be some interactions lurking around). What is worse, the problem persists that if a given interaction is not observed on a given day, this could be either because it was impossible that day despite being otherwise feasible (P($X|L=1$=0), because interaction between the two did occur but could not be observed (P($D|X$=1,L=1)=0), or from any combination of the two. From a conceptual perspective, this approach will thus fail to satisfactorily distinguish between sources of uncertainty. An interaction not observed on day $t$ may have gone undetected either because there were constraints on the process (i.e. the probability $\lambda$ was really zero) or because of chance variation in the realized incidence of an event associated with a finite probability. Most importantly, if two species are never observed co-occurring during several days of sampling then we have learned nothing about their probability of interacting if they should ever co-occur. In other words, there is no information without co-occurance.

@@

  As an alternative to abandoning empirical networks or continuing to ignore the uncertainty inherent in undetected observations, we propose that some insight regarding the detectability of interactions between species not found co-occurring may be gained from data on other species between which interactions are known to be feasible. If we use this information obtain the joint probability of detection and process (e.g. to glean information regarding at how many sites the species are found interacting across the entire dataset), and assuming that most of this pattern is due to variation in detection rate, then insights into the incidence of interactions between species may be gleaned from the accumulation rate of positive interactions at sites were interaction is possible (i.e. with at least one observation of a positive interaction) [this reasoning to be further formalized]

@@

      \begin{equation}
        \underbrace{P(\theta|X,n)}_{Posterior} = \frac{\overbrace{P(X|\theta,n)}^{Likelihood}\overbrace{P(\theta)}^{Prior}}{\underbrace{P(X|N)}_{Normalizer}} .
        \label{posterior}
      \end{equation}

@@

     In some cases it can be complicated to compute the normalizer, but fortunately in our case an analytical solution exists. The beta-binomial distribution is a conjugate distribution of the binomial distribution. This allows us to analytically compute the posterior distribution of a binomial model with a beta prior distribution. We can re-write the posterior distribution of $\theta$ as:

@@

      \begin{figure}[ht]
        \label{Salix_pdfs}
        \caption{Using prior distribution based on the \emph{Salix}-galler and galler-parasitoid networks in~\citet{Barbour2016}, we can calculate posterior distributions for the probability of interaction ($\lambda$) between two species that have not yet been observed interacting. Here we show posterior distributions for each $\lambda$ in each network ranging from the prior distribution (N=0 observed co-occurances) to the distribution obtained when the pair of species has been observed co-occurring 100 times. The distribution narrows and approaches zero as the sample size increases. \textbf{A)} The posterior distributions for the \emph{Salix}-galler component are always wider and farther from zero than those for \textbf{B)} the galler-parasitoid component.}
        \includegraphics*[width=.8\textwidth]{figures/Salix_Galler_pdfs_increasing_N.eps}
        \end{figure}
        [[Maybe I was using the int_probs from the other example?]]

@@

          In absence of any external information, an uniformative prior is the most conservative hypothesis for the distribution of $\theta$. The beta distribution is in this case a uniform distribution  for $Beta(\alpha=1,\beta=1)$.

@@

          [[Should we still have this? At the workshop we talked about how interaction probabilities depend on the degrees of both species and so I'm using predicted interaction probabilities (degreeA*degreeB) in the galler example. A writeup for that process follows the original text:]]
          The degree of a node in a network is defined as the number of interactions it has to other nodes. The degree distribution of a network is then the probability distribution of these degrees over the whole network. Connectance is an average property for a complete given network, and it represents the expected degree (expected number of interactions per species) standardized by the number of species in the network. The standardized degree could therefore be interpreted as an interaction probability. It is consequently possible to use the degree distribution to inform the prior distribution. The degree distribution could come from several networks, from a similar network (e.g. a known network at slightly different location) or from the network of interest if interaction probabilities for some species are already documented. The latter approach allows us to apply information from known, abundant species to the rarest species for which interactions are less frequently documented. The procedure for the estimation of the hyper parameters follows the same approach as described above for connectance except that each measurement is at the individual species level instead of the network level.

          If our focal network describes a similar system to that in a known network, we can use the distribution of interaction probabilities in that network to inform our prior. The probability of any interaction $L_ij$ depends on the degrees of species $i$ and $j$. The degree of a node is defined as the number of interactions it has with other nodes. Connectance, as the average number of links per species in the whole network, is thus also the expected degree for any given species normalized by the number of species in the network. Using normalized degrees $\Delta_i$ and $\Delta_j$ (i.e., degrees divided by the number of species in the network), we can obtain the probability of interaction $L_ij$=$\Delta_i\times\Delta_j$. The distribution of these interaction probabilities derived from a similar network or a suite of networks can be used to establish a prior distribution before any data from the focal network are collected. We could also use this approach to build a prior distribution using known interaction probabilities for some species in the focal network. The latter approach allows us to apply information from known, abundant species to the rarest species for which interactions are less frequently documented. The procedure for the estimation of the hyper parameters follows the same approach as described above for connectance except that each measurement is at the individual interaction level instead of the network level.

@@

          It is possible to estimate the probability of interaction between a pair of species if there is knowledge of their traits and some functions relating trait matching to interaction probability~\citep{Morales-Castilla2015}. There are several techniques available to perform this inferrence of interaction probability~\citep{} (REFS). Note that in this case the prior might not be beta distribution and numerical methods might be required to compute the posterior distribution.

@@

  %Perhaps we want to also have a trait matching example in addition to the degree distribution example below?
  % Are degree distributions valid priors for interaction probabilities in a unipartite web? For the Salix-gall dataset we're using interaction probabilities as priors (since the degree of both Salix and galler matter)... maybe we should rephrase this to predict the degree of species $i$ based on the degrees of 10 other species in the network?
  The Bayesian framework can be illustrated with a simple quantiative example. Suppose we have $n = 10$ observations of co-occurrence between species $i$ and species $j$ in a given time interval and area, and $X = 3$ observations of interactions. The maximum likelihood estimate of the interaction probability is simply $\theta_{MLE} = 3/10 = 0.3$. 
  % We should probabaly have a value  on the interaction probablitity that differs from the average of the degree distrbution
   Now consider we know that species $i$ is known to interact with 10 other species, which have the following normalized degrees:

@@

    If the network has 20 species total, then species $i$ has a normalized degree of 0.5 and
    we can use this to obtain the following set of interaction probabilities for species $i$ and each of its interaction partners:

@@

    It is fairly straigthforward to compute most of network metrics when the different $\lambda$ of the adjacency matrix are known, without variance~\citep{Poisot2016}. Several of these metrics derive directly from quantitative indices of network structure which are equivalent to $\lambda$; the remainder, originally defined for binary networks, can be adjusted to account for interaction probabilities between zero and one. It is not as easy, however, to understand how the uncertainty in these estimated interaction probabilities influences network metrics. Computation of these metrics involves non-linear functions and therefore, because of Jensen's inequality (which states that the average of non-linear function of a stochastic variable differs from the function of the average of that variable), any uncertainty in $\lambda$s could bias both the mean and variance of a network metric. For instance, assume that among all pairs of species with few observed co-occurances there are many pairs which truly interact.    If we begin with no observed links for these pairs, then correcting for these false absences will inflate the number of links and consequently influence metrics such as connectance, nestedness, and degree distribution. While we can, as described above, calculate these network metrics using fixed estimates for $\lambda$ no analytical solution currently exists for uncertain $\lambda$. Instead we must rely on computer simulations. In the following section, we will provide an example with the Salix dataset and show how the uncertainty arising from both absences of co-occurrences and false absences of interactions impact these metrics. We find uncertainty influences both the average and the variance of network metrics, putting in perspective traditional network comparison studies where these metrics are assumed fixed entities.

@@

      In a strict Bayesian framework, we wish to use a prior distribution that does not rely on any information from the study at hand. To that end, we use data from another well-described \emph{Salix} galler-parasitoid system~\citep{Barbour2016,Barbour2016Dryad}. Note that this study used several genotypes of \emph{S. hookeriana} rather than different \emph{Salix} species. We estimated frequencies of  \emph{S. hookeriana} genotype-galler interactions based on the normalized degree of each species in each network component (see \emph{Appendix S1} for details and code).

      Using this prior distribution of galler-\emph{Salix} interaction frequencies, we can obtain parameters for the prior distribution of $\lambda$ in our dataset. Specifically, we obtain $\alpha$=2.507217, $\beta$=1.888079. For the galler-parasitoid component of the network, the corresponding prior parameters are $\alpha$=1.341479, $\beta$=9.487966.
      We are then able to use these priors to estimate the posterior distriution of interaction probabilities given the additional information in our dataset.

@@

      For species where no co-occurances were observed, we can 
      calculate the MLE estimates for the mean and variance of 
      $\lambda_{ij}$ directly from this prior following 
      equations~\ref{mean} and~\ref{variance} 
      (see \emph{Appendix S2} for R implementation). The prior 
      distribution we obtain based on~\citet{Barbour2016,Barbour2016Dryad} is:
      $\hat{\lambda_{ij}}$=0.570, var($\lambda$)=0.056 for the \emph{Salix}-galler network and $\hat{\lambda_{ij}}$=0.124, var($\lambda$)=0.010 for the galler-parasitoid network.
      This is consistent with the prior data, where the average 
      interaction probability based on the degree distributions of Salix genotypes and gallers was 0.562 and that for gallers and parasitoids was 0.122. [[What's the connectance of our dataset, if we call greys 0?]] This probability does, however, seem rather high for our dataset. This discrepancy may reflect the differences in scale between a network containing different genotypes of a single \emph{Salix} species and four gallers from a single family and a network containing many \emph{Salix} species and gallers from many families as is the case in our dataset.

      For a pair of species where some co-occurances were observed, we can update the prior distribution with these data. If we consider only pairs of species which were observed to co-occur but not to interact, $k_{ij}$ is always 0 and only $n_{ij}$ will vary between species pairs. Thus $\alpha'$=$\alpha$ and $\beta'$=$\beta + n$. At the most extreme case, for a pair of species which co-occurred at all 374 sites and was never observed to interact, our distributions would become:

@@

      With species $i$ and $j$ co-occurring at all 374 sites and never being observed to interact, the maximum likelihood expectation for the probability of an interaction $ij$ is very close to 0 and the variance about this estimate is very narrow. In most cases, however, species $i$ and $j$ did not co-occur at so many sites and our posterior mean and variance retain some influence of the prior. Nevertheless, as the number of observed co-occurances increases, the posterior mean becomes closer to zero and the posterior variance decreases (Fig.~\ref{Salix_pdfs}).

@@

      As well as obtaining posterior means and variances, it is possible to calculate a 95\% credible interval about the posterior mean. This is again straightforward to do in R (see the function ``credible interval" in \emph{Appendix S2}). Calculating the 95\% credible intervals for species pairs with 0 to 50 observed co-occurrances without any observed interaction (as in Fig.~\ref{Salix_pdfs}), we see that the interval narrows dramatically as the number of observations increases. In no case, however, does the confidence interval include 0. Even for a pair of species which appeared together in all 374 sites but were never observed interacting, the confidence interval (0.001, 0.018) comes close to, but does not include zero (Table~\ref{interval_table}).

@@

>>>>

    The main objective of this study is to develop analytical tools that could be used to better represent the uncertainty in the estimation of pairwise interaction probability. We start with the description of a new Bayesian approach to estimate ecological interactions and of a related R package. [[We first illustrate the framework with a simple quantitative example. Next, we use the framework to provide analytical criteria for improving the sampling of rare interactions. Finally, we discuss the implications of uncertainty in the estimation of pairwise interactions on the properties of ecological networks.]][[Order has probably changed.]] Through these efforts, we demonstrate both the utility of our approach and the importance of acknowledging the uncertainty inherent in ecological networks.
    % List of sections may not be accurate any more. Update when paper reaches its final form.

@@

THIS SECTIONS BUILDS ON THE SALIX DATA TO ILLUSTRATE THE GAPS THERE ARE IN THE METHOD. DESPITE THE INTENSITY OF THE SAMPLING, THERE ARE SEVERAL HOLES.[[Still necessary? If so, we should flesh this out into a paragraph.]]

\section*{A naive quantification of interaction probability and uncertainty}

@@

  As a start, we will consider how we could naively quantify interaction probability and its associated uncertainty for an interaction that has not yet been observed. We consider that a pair of species have been observed $n$ times co-occurring at one location, and the number of times they have been observed interacting is $k = 0$. We aim to evaluate the uncertainty of this interaction. The occurrence of an interaction is a Bernoulli trial. Consequently, the number of success $k$ over $n$ trials will follow a binomial distribution:

@@

  The variance of a Bernoulli experiment is simply $n\lambda$(1-$\lambda$). It is important to remember this variance reports the variability of the number of success $k$ for $n$ trials and is not the variance associated with the estimation of $\lambda$.
  Given this variance, it is possible to compute the confidence interval for the MLE of $\lambda$ using any of several methods, including the \emph{Wilson score interval}, the \emph{Clopper-Pearson interval}, and the \emph{Agresti-Coull Interval}. Finding this estimate is therefore quite straightforward, but it nonetheless has two drawbacks. First, $\lambda$ is a random variable with a given uncertainty and whose distribution is unknown, it is not a single point estimate. This means that even if $k = 0$ in a given sample, this does not necessarily imply that the two species will never interact. It means that no interaction is the most likely situation but there is nonetheless a significant chance that the two species could interaction. In contrast with the situation where $k>0$, when we are sure that an interaction is feasible, 
  % (in which case, the ratio $k/n$ is smaller than $n$ because of detection and process uncertainty),  % This parenthetical makes the sentence very long. Is it necessary?
  we are not sure that absence of interactions is caused by $L=0$. Second, and most importantly, the number of trials $n$ might be very low in many instances (and some pairs of species might even not be documented once). Consequently, the uncertainty in the estimation of $\lambda$ might be considerable. We illustrated in Fig. X how the estimate of $\lambda$ could vary with sampling size using the Clopper-Pearson interval. It shows that the 95\% confidence interval requires a significant amount of sampling in order to be sure that species are not interacting. 
  % Is the Clopper-Pearson interval the one in Dom's code snippets? If so, this is Fig.~\ref{Salix_cdfs}

@@

  What does this question mean exactly? We consider that there are three nested levels of uncertainty making the occurrence of any interaction a stochastic process:

@@

    It could be that the interaction is feasible, i.e. $L=1$, but that it does not occur at a given location or moment in time because there are local constraints preventing it from occurring. We define the realisation of the interaction process $X$ given that the interaction is feasible as a stochastic process with associated probability $P(X|L)=\chi$. This phenomenon of interaction contingencies is usually not considered in network studies, but there is a rich literature in community ecology about the contingencies of interactions~\citep{}. We imagine for instance that an interaction between a gall and a parasitoid might not be recorded at a location because it was the wrong time of the year and the gall had not yet formed (phenological constraints), a late frost killed the larvae (an abiotic environmental constraint), another parasitoid species competitively excluded the parasitoid species of interest (a biotic environmental constraint) or simply the parasitoid is too rare to interact with all of its viable hosts (an abundance constraint).

@@

    Lastly, measurement errors are always source of uncertainty in the observation of ecological processes. We define the detection of an interaction $D$, given an interaction is feasible and occurs in the local conditions, as a stochastic process with associated probability $P(D|X,L)=\delta$. Detection failure could happen for several reasons, for instance if the rearing of a parasitoid fails with inappropriate lab conditions or because of species mis-identification. Some sources of detection error could be minimised with appropriate sampling effort ($\delta$ will converge to one with increasing number of collected galls), but other sources are often difficult to reduce (e.g. the occurrence of cryptic species might require molecular analysis for appropriate taxonomic identification).

@@

  For the ecologist working on collecting network data, and on later interpreting these data, the considerations above raise a major challenge: how may we infer whether unobserved interactions went undetected due to sampling, or whether they truly do not occur. How then may we refine our sampling approaches to reduce uncertainties, and to gain insights into the impact of multiple processes on field observations? Importantly, some sources can be minimised with appropriate sampling design and efforts, while other sources are difficult or impossible to reduce – since they are generated by chance variation created by the very process in which we are interested.

  The basic information that we are collecting consists of matrices of interaction partners. Our count of interactions between species A and species B reflects uncertainty with respect to whether two species interact when co-occurring at a given site, chance variation in whether the interaction is realised, and uncertainty with respect to whether it is detected when realised. In this case, what can we do to pin down any of these different types of variation?

@@

  \begin{figure}
    \caption{\textbf{A-B)} Most pairs of \emph{Salix} and galler or galler and parasitoid species were never observed co-occurring despite the high levels of replication in our example dataset. For those that were observed together at least once, the number of observed co-occurrences was generally small (\textless10). Here we show a histogram of the number of pairs of species observed co-occurring at least once. 3986 pairs of species were never observed co-occurring: these pairs are omitted from the histogram. \textbf{C-D)} Most pairs of species that were observed at the same site were never observed interacting. Here we show a histogram of the number of observed interactions within pairs of co-occurring species. \textbf{E-F)} Here we show, for each species pair, the number of observed interactions plotted against the number of observed co-occurrences. Most pairs either are never observed interacting or interact almost every time they co-occur. The red, dashed line indicates a 1:1 relationship between interactions and co-occurrences.}
    \label{histograms}
    \includegraphics*[width=.8\textwidth]{figures/Salix_Galler_histogram.eps}
    \end{figure}

@@

  As a solution towards deciding whether unobserved interactions were undetected due to sampling, or whether they truly do not occur,~\citet{Weinstein2017} used daily repeated sampling rounds to estimate the daily probability of detecting a hummingbird interaction, and to thereby model the observation and process mechanisms~\citep{Weinstein2017}. While conceptually attractive, this approach is unsuitable for interactions occurring over longer time scales (e.g., associations between hosts and parasitoids with a single generation per year), or very rare interactions (since no matter how frequently and extensively we sample, there will still be some interactions lurking around). What is worse, the problem persists that if a given interaction is not observed on a given day, this could be either because it was impossible that day despite being otherwise feasible (P($X|L=1$=0), because interaction between the two did occur but could not be observed (P($D|X$=1,L=1)=0), or from any combination of the two. From a conceptual perspective, this approach will thus fail to satisfactorily distinguish between sources of uncertainty. An interaction not observed on day $t$ may have gone undetected either because there were constraints on the process (i.e. the probability $\lambda$ was really zero) or because of chance variation in the realised incidence of an event associated with a finite probability. Most importantly, if two species are never observed co-occurring during several days of sampling then we have learned nothing about their probability of interacting if they should ever co-occur. In other words, there is no information without co-occurrence.

@@

  As an alternative to abandoning empirical networks or continuing to ignore the uncertainty inherent in undetected observations, we propose that some insight regarding the detectability of interactions between species not found co-occurring may be gained from data on other species between which interactions are known to be feasible. If we use this information obtain the joint probability of detection and process (e.g. to glean information regarding at how many sites the species are found interacting across the entire dataset), and assuming that most of this pattern is due to variation in detection rate, then insights into the incidence of interactions between species may be gleaned from the accumulation rate of positive interactions at sites were interaction is possible (i.e. with at least one observation of a positive interaction) [this reasoning to be further formalised]

@@

      \begin{equation}
        \underbrace{P(\theta|X,n)}_{Posterior} = \frac{\overbrace{P(X|\theta,n)}^{Likelihood}\overbrace{P(\theta)}^{Prior}}{\underbrace{P(X|N)}_{Normaliser}} .
        \label{posterior}
      \end{equation}

@@

     In some cases it can be complicated to compute the normaliser, but fortunately in our case an analytical solution exists. The beta-binomial distribution is a conjugate distribution of the binomial distribution. This allows us to analytically compute the posterior distribution of a binomial model with a beta prior distribution. We can re-write the posterior distribution of $\theta$ as:

@@

      \begin{figure}[ht]
        \label{Salix_pdfs}
        \caption{Using prior distribution based on the \emph{Salix}-galler and galler-parasitoid networks in~\citet{Barbour2016}, we can calculate posterior distributions for the probability of interaction ($\lambda$) between two species that have not yet been observed interacting. Here we show posterior distributions for each $\lambda$ in each network ranging from the prior distribution (N=0 observed co-occurrence) to the distribution obtained when the pair of species has been observed co-occurring 100 times. The distribution narrows and approaches zero as the sample size increases. \textbf{A)} The posterior distributions for the \emph{Salix}-galler component are always wider and farther from zero than those for \textbf{B)} the galler-parasitoid component.}
        \includegraphics*[width=.8\textwidth]{figures/Salix_Galler_pdfs_increasing_N.eps}
        \end{figure}

@@

          In absence of any external information, an uninformative prior is the most conservative hypothesis for the distribution of $\theta$. The beta distribution is in this case a uniform distribution  for $Beta(\alpha=1,\beta=1)$.

@@

          [[Should we still have this? At the workshop we talked about how interaction probabilities depend on the degrees of both species and so I'm using predicted interaction probabilities (degreeA*degreeB) in the galler example. A writeup for that process follows the original text:]]
          The degree of a node in a network is defined as the number of interactions it has to other nodes. The degree distribution of a network is then the probability distribution of these degrees over the whole network. Connectance is an average property for a complete given network, and it represents the expected degree (expected number of interactions per species) standardised by the number of species in the network. The standardised degree could therefore be interpreted as an interaction probability. It is consequently possible to use the degree distribution to inform the prior distribution. The degree distribution could come from several networks, from a similar network (e.g. a known network at slightly different location) or from the network of interest if interaction probabilities for some species are already documented. The latter approach allows us to apply information from known, abundant species to the rarest species for which interactions are less frequently documented. The procedure for the estimation of the hyper parameters follows the same approach as described above for connectance except that each measurement is at the individual species level instead of the network level.

          If our focal network describes a similar system to that in a known network, we can use the distribution of interaction probabilities in that network to inform our prior. The probability of any interaction $L_ij$ depends on the degrees of species $i$ and $j$. The degree of a node is defined as the number of interactions it has with other nodes. Connectance, as the average number of links per species in the whole network, is thus also the expected degree for any given species normalised by the number of species in the network. Using normalised degrees $\Delta_i$ and $\Delta_j$ (i.e., degrees divided by the number of species in the network), we can obtain the probability of interaction $L_ij$=$\Delta_i\times\Delta_j$. The distribution of these interaction probabilities derived from a similar network or a suite of networks can be used to establish a prior distribution before any data from the focal network are collected. We could also use this approach to build a prior distribution using known interaction probabilities for some species in the focal network. The latter approach allows us to apply information from known, abundant species to the rarest species for which interactions are less frequently documented. The procedure for the estimation of the hyper parameters follows the same approach as described above for connectance except that each measurement is at the individual interaction level instead of the network level.

@@

          It is possible to estimate the probability of interaction between a pair of species if there is knowledge of their traits and some functions relating trait matching to interaction probability~\citep{Morales-Castilla2015}. There are several techniques available to perform this inference of interaction probability~\citep{} (REFS). Note that in this case the prior might not be beta distribution and numerical methods might be required to compute the posterior distribution.

@@

  %Perhaps we want to also have a trait matching example in addition to the degree distribution example below?
  % Are degree distributions valid priors for interaction probabilities in a unipartite web? For the Salix-gall dataset we're using interaction probabilities as priors (since the degree of both Salix and galler matter)... maybe we should rephrase this to predict the degree of species $i$ based on the degrees of 10 other species in the network?
  The Bayesian framework can be illustrated with a simple quantitative example. Suppose we have $n = 10$ observations of co-occurrence between species $i$ and species $j$ in a given time interval and area, and $X = 3$ observations of interactions. The maximum likelihood estimate of the interaction probability is simply $\theta_{MLE} = 3/10 = 0.3$. 
  % We should probabaly have a value  on the interaction probablitity that differs from the average of the degree distrbution
   Now consider we know that species $i$ is known to interact with 10 other species, which have the following normalised degrees:

@@

    If the network has 20 species total, then species $i$ has a normalised degree of 0.5 and
    we can use this to obtain the following set of interaction probabilities for species $i$ and each of its interaction partners:

@@

    It is fairly straightforward to compute most of network metrics when the different $\lambda$ of the adjacency matrix are known, without variance~\citep{Poisot2016}. Several of these metrics derive directly from quantitative indices of network structure which are equivalent to $\lambda$; the remainder, originally defined for binary networks, can be adjusted to account for interaction probabilities between zero and one. It is not as easy, however, to understand how the uncertainty in these estimated interaction probabilities influences network metrics. Computation of these metrics involves non-linear functions and therefore, because of Jensen's inequality (which states that the average of non-linear function of a stochastic variable differs from the function of the average of that variable), any uncertainty in $\lambda$s could bias both the mean and variance of a network metric. For instance, assume that among all pairs of species with few observed co-occurrence there are many pairs which truly interact.    If we begin with no observed links for these pairs, then correcting for these false absences will inflate the number of links and consequently influence metrics such as connectance, nestedness, and degree distribution. While we can, as described above, calculate these network metrics using fixed estimates for $\lambda$ no analytical solution currently exists for uncertain $\lambda$. Instead we must rely on computer simulations. In the following section, we will provide an example with the Salix dataset and show how the uncertainty arising from both absences of co-occurrences and false absences of interactions impact these metrics. We find uncertainty influences both the average and the variance of network metrics, putting in perspective traditional network comparison studies where these metrics are assumed fixed entities.

@@

      In a strict Bayesian framework, we wish to use a prior distribution that does not rely on any information from the study at hand. To that end, we use data from another well-described \emph{Salix} galler-parasitoid system~\citep{Barbour2016,Barbour2016Dryad}. Note that this study used several genotypes of \emph{S. hookeriana} rather than different \emph{Salix} species. We estimated frequencies of  \emph{S. hookeriana} genotype-galler interactions based on the normalised degree of each species in each network component (see \emph{Appendix S1} for details and code).

      Using this prior distribution of galler-\emph{Salix} interaction frequencies, we can obtain parameters for the prior distribution of $\lambda$ in our dataset. Specifically, we obtain $\alpha$=2.507217, $\beta$=1.888079. For the galler-parasitoid component of the network, the corresponding prior parameters are $\alpha$=1.341479, $\beta$=9.487966.
      We are then able to use these priors to estimate the posterior distribution of interaction probabilities given the additional information in our dataset.

@@

      For species where no co-occurrence were observed, we can 
      calculate the MLE estimates for the mean and variance of 
      $\lambda_{ij}$ directly from this prior following 
      equations~\ref{mean} and~\ref{variance} 
      (see \emph{Appendix S2} for R implementation). The prior 
      distribution we obtain based on~\citet{Barbour2016,Barbour2016Dryad} is:
      $\hat{\lambda_{ij}}$=0.570, var($\lambda$)=0.056 for the \emph{Salix}-galler network and $\hat{\lambda_{ij}}$=0.124, var($\lambda$)=0.010 for the galler-parasitoid network.
      This is consistent with the prior data, where the average 
      interaction probability based on the degree distributions of Salix genotypes and gallers was 0.562 and that for gallers and parasitoids was 0.122. These probabilities are much higher than those in our dataset (0.028 and 0.078 for the \emph{Salix}-galler and galler-parasitoid networks, respectively). This discrepancy may reflect the differences in scale between a network containing different genotypes of a single \emph{Salix} species and four gallers from a single family and a network containing many \emph{Salix} species and gallers from many families as is the case in our dataset.

      For a pair of species where some co-occurrence were observed, we can update the prior distribution with these data. If we consider only pairs of species which were observed to co-occur but not to interact, $k_{ij}$ is always 0 and only $n_{ij}$ will vary between species pairs. Thus $\alpha'$=$\alpha$ and $\beta'$=$\beta + n$. At the most extreme case, for a pair of species which co-occurred at all 374 sites and was never observed to interact, our distributions would become:

@@

      With species $i$ and $j$ co-occurring at all 374 sites and never being observed to interact, the maximum likelihood expectation for the probability of an interaction $ij$ is very close to 0 and the variance about this estimate is very narrow. In most cases, however, species $i$ and $j$ did not co-occur at so many sites and our posterior mean and variance retain some influence of the prior. Nevertheless, as the number of observed co-occurrence increases, the posterior mean becomes closer to zero and the posterior variance decreases (Fig.~\ref{Salix_pdfs}).

@@

      As well as obtaining posterior means and variances, it is possible to calculate a 95\% credible interval about the posterior mean. This is again straightforward to do in R (see the function ``credible interval" in \emph{Appendix S2}). Calculating the 95\% credible intervals for species pairs with 0 to 50 observed co-occurrences without any observed interaction (as in Fig.~\ref{Salix_pdfs}), we see that the interval narrows dramatically as the number of observations increases. In no case, however, does the confidence interval include 0. Even for a pair of species which appeared together in all 374 sites but were never observed interacting, the confidence interval (0.001, 0.018) comes close to, but does not include zero (Table~\ref{interval_table}).

@@

    Note that the posterior webs for the Salix-galler network all had much higher connectance than the original, observed web (C=0.028 for the observed web, Fig.~\ref{posterior_webs}). Likewise, the number of links per \emph{Salix} and galler species in the observed web (2.71 and 1.47, respectively) were much lower than those predicted in the posterior webs, even when filtering with a detection probability of only 50\%. There are two possible explanations for this discrepancy: either the~\citet{Barbour2016} data do no give an appropriate prior for our dataset, or the true detection probability for links between \emph{Salix} and galler species is much less than 50\%. As the scale of the two datasets is quite different (genotypes of a single \emph{Salix} species in~\citet{Barbour2016}, various \emph{Salix} species in our dataset), we suspect the former is more likely.
